# 2021-10-08 - Additional math tasks

## Overview

```{r}
experiment_slug <- "2021-10-08-math-tasks"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(scales)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

# Mark factors
exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tournament",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

exp_summary_data$NUM_POPS <- factor(
  exp_summary_data$NUM_POPS,
  levels=c(
    "24",
    "48",
    "96"
  )
)

exp_summary_data$AVIDAGP_ENV_FILE <- factor(
  exp_summary_data$AVIDAGP_ENV_FILE,
  levels=c(
    "environment-small.json",
    "environment-big.json"
  ),
  labels=c(
    "ENV-SMALL",
    "ENV-BIG"
  )
)

exp_summary_data$UPDATES_PER_EPOCH <- as.factor(
  exp_summary_data$UPDATES_PER_EPOCH
)

exp_summary_data$POPULATION_SAMPLING_SIZE <- as.factor(
  exp_summary_data$POPULATION_SAMPLING_SIZE
)

exp_summary_data$SAMPLE_SIZE <- exp_summary_data$POPULATION_SAMPLING_SIZE
exp_summary_data$ENV <- exp_summary_data$AVIDAGP_ENV_FILE
exp_summary_data$U_PER_E <- exp_summary_data$UPDATES_PER_EPOCH
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Paired"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)
```

## Average number of organisms

Average number of organisms per world at the end of a run.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_num_orgs,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_num_orgs.pdf")
)
```

## Average generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_gens,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_gens.pdf"),
  width=10,
  height=15
)
```

## Performance

### CPU cycles per replication

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_cpu_cycles_per_replication,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_cpu_cycles_per_replication.pdf")
)
```

### Max aggregate performance

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=max_aggregate_score,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="max_aggregate_score (log scale)",
    trans=pseudo_log_trans(sigma = 1, base = 10),
    breaks=c(0, 10, 100, 1000, 10000, 100000)
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_aggregate_score.pdf")
)
```

### Max trait coverage

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=max_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_trait_coverage.pdf")
)
```

### Total trait coverage

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=total_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "total_trait_coverage.pdf")
)
```

## Selection

### Average number of unique populations selected

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_unique_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_unique_selected.pdf")
)
```

### Average entropy of selection ids

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_entropy_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_entropy_selected.pdf")
)
```

## 2021-11-03 Lab meeting

```{r}
ggplot(
    filter(
      exp_summary_data,
      EPOCHS=="1000" & U_PER_E=="300" & ENV=="ENV-SMALL"
    ),
    aes(
      x=SELECTION_METHOD,
      y=max_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "2021-11-03-max_trait_coverage.png")
)
```

```{r}
ggplot(
    filter(
      exp_summary_data,
      EPOCHS=="1000" & U_PER_E=="300" & ENV=="ENV-SMALL"
    ),    aes(
      x=SELECTION_METHOD,
      y=avg_cpu_cycles_per_replication,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    ENV~EPOCHS+U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "2021-11-03-avg_cpu_cycles_per_replication.png")
)
```




