# Directed digital evolution experiment

Supplemental information and data analyses for the initial (baseline) directed evolution modeling experiment.

## Overview

```{r}
experiment_slug <- "2021-11-11-selection"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Load experiment summary data.

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "elite-10",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "elite-10",
    "tourn",
    "lex",
    "nde",
    "random",
    "none"
  )
)

```

Load time series data.

```{r}
times_series_data_loc <- paste0(working_directory,"data/evaluation_time_series.csv")
times_series_data <- read.csv(times_series_data_loc, na.strings="NONE")

# Specify experimental condition for each datum.
times_series_data$SELECTION_METHOD <- factor(
  times_series_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "elite-10",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "elite-10",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)
```

Load task coverage per population data.

```{r}
task_coverage_per_pop_data_loc <- paste0(working_directory,"data/max_coverage_per_pop_cnt.csv")
task_coverage_per_pop_data <- read.csv(task_coverage_per_pop_data_loc, na.strings="NONE")

# Specify experimental condition for each datum.
task_coverage_per_pop_data$SELECTION_METHOD <- factor(
  task_coverage_per_pop_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "elite-10",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "elite-10",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Set2"
alpha <- 0.05
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)

p_label <- function(p_value) {
  threshold = 0.0001
  if (p_value < threshold) {
    return(paste0("p < ", threshold))
  } else {
    return(paste0("p = ", p_value))
  }
}


selection_method_breaks <- c("elite", "elite-10", "tourn", "lex", "nde", "random", "none")
selection_method_labels <- c("ELITE", "TOP-10", "TOURN", "LEX", "NDE", "RAND", "NONE")
```

## Best single-population task coverage

Best single-population task coverage at the end of the experiment.

```{r}
max_trait_cov_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=max_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5,18.5),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )

max_trait_cov_fig
ggsave(
  plot=max_trait_cov_fig,
  paste0(plot_directory, "max_trait_coverage.pdf")
)
```

Statistical results:

```{r}
kruskal.test(
  formula=max_trait_coverage~SELECTION_METHOD,
  data=exp_summary_data
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_summary_data$max_trait_coverage,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

### Best single-population task coverage time series

To speed up graphing, we plot a low-resolution version of the time series.

```{r}
max_trait_cov_ot_fig <-
  ggplot(
    filter(times_series_data, (updates_elapsed%%10000)==0 | updates_elapsed==1),
    aes(
      x=updates_elapsed,
      y=max_trait_coverage,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_x_continuous(
    name="Updates elapsed"
  ) +
  scale_y_continuous(
    name="Task Coverage",
    breaks=seq(0,18,2)
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="bottom"
  )
max_trait_cov_ot_fig
ggsave(
  plot=max_trait_cov_ot_fig,
  filename=paste0(plot_directory, "2021-11-11-best-pop-task-cov-ts.png"),
  width=10,
  height=6
)
```

## Metapopulation task coverage

Metapopulation task coverage at the end of the experiment.

```{r}
total_trait_cov_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=total_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5,18),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )

total_trait_cov_fig
ggsave(
  plot=total_trait_cov_fig,
  paste0(plot_directory, "2021-11-11-metapop-task-cov.pdf")
)
```

Statistical results:

```{r}
kruskal.test(
  formula=total_trait_coverage~SELECTION_METHOD,
  data=exp_summary_data
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_summary_data$total_trait_coverage,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)

# All sig p<0.5, except tourn vs none
```

### Metapopulation task coverage time series

To speed up plotting, we graph a low-resolution version of this time series.

```{r}
metapop_task_cov_ot_fig <-
  ggplot(
    filter(times_series_data, (updates_elapsed%%10000)==0 | updates_elapsed==1),
    aes(
      x=updates_elapsed,
      y=total_trait_coverage,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_x_continuous(
    name="Updates elapsed"
  ) +
  scale_y_continuous(
    name="Task Coverage",
    breaks=seq(0,18,2)
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="bottom"
  )
metapop_task_cov_ot_fig
ggsave(
  plot=metapop_task_cov_ot_fig,
  filename=paste0(plot_directory, "2021-11-11-metapop-task-cov-ts.png"),
  width=10,
  height=6
)
```

## Metapopulation task profile diversity

We measured the "phenotypic" diversity within evolved metapopulations in three ways:

1. the number of task profiles (richness)
2. the spread of task profiles as the average cosine distance from the centroid profile
3. the Shannon entropy of task profiles

### Number of different task profiles

```{r}
num_pop_task_profiles_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=num_pop_trait_profiles,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="# Different Task Profiles"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )
num_pop_task_profiles_fig
ggsave(
  plot=num_pop_task_profiles_fig,
  paste0(plot_directory, "2021-11-11-num-task-profiles.pdf")
)
```

Statistical results:

```{r}
kruskal.test(
  formula=num_pop_trait_profiles~SELECTION_METHOD,
  data=exp_summary_data
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_summary_data$num_pop_trait_profiles,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

#### Number of different task profiles over time

```{r}
num_task_profiles_ot_fig <-
  ggplot(
    filter(times_series_data, (updates_elapsed%%10000)==0 | updates_elapsed==1),
    aes(
      x=updates_elapsed,
      y=num_pop_trait_profiles,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_x_continuous(
    name="Updates elapsed"
  ) +
  scale_y_continuous(
    name="# Different Task Profiles"
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="bottom"
  )
num_task_profiles_ot_fig
ggsave(
  num_task_profiles_ot_fig,
  filename=paste0(plot_directory, "2021-11-11-num-task-profiles-ts.png"),
  width=10,
  height=6
)
```

### Task profile spread

```{r}
task_profile_spread_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_cosine_dist_from_centroid,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Avg. Task Spread"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )
task_profile_spread_fig
ggsave(
  plot=task_profile_spread_fig,
  paste0(plot_directory, "2021-11-11-task-profile-spread.pdf")
)
```

Statistical results:

```{r}
kruskal.test(
  formula=avg_cosine_dist_from_centroid~SELECTION_METHOD,
  data=exp_summary_data
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_summary_data$avg_cosine_dist_from_centroid,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

### Task profile entropy

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=pop_trait_profile_entropy,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Shannon entropy of task profiles"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "2021-11-11-task-profile-entropy.pdf")
)
```

Statistical results:

```{r}
kruskal.test(
  formula=pop_trait_profile_entropy~SELECTION_METHOD,
  data=exp_summary_data
)

pairwise.wilcox.test(
  x=exp_summary_data$pop_trait_profile_entropy,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

#### Task entropy over time

```{r}
ggplot(
    filter(times_series_data, (updates_elapsed%%10000)==0 | updates_elapsed==1),
    aes(
      x=updates_elapsed,
      y=pop_trait_profile_entropy,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_x_continuous(
    name="Updates elapsed"
  ) +
  scale_y_continuous(
    name="Task profile entropy"
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  filename=paste0(plot_directory, "2021-11-11-pop-trait-profile-entropy-ts.png"),
  width=10,
  height=6
)
```

## Task coverage per N populations

We analyzed the (maximum) number of tasks added to metapopulation task coverage for a given number (N) of member populations considered.
That is, for each N, we solved the maximum set coverage problem for task coverage: what is the maximum number of tasks that can be covered given N populations from this metapopulation?

```{r}
ggplot(
    task_coverage_per_pop_data,
    aes(
      x=n_pops,
      y=max_tasks_covered,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    name="Maximum task coverage"
  ) +
  scale_x_continuous(
    name="Number of populations",
    limits=c(0, 15)
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  paste0(plot_directory, "2021-11-11-task-cov-per-n-pops.pdf"),
  width=10,
  height=6
)
```

```{r}
ggplot(
    filter(
      task_coverage_per_pop_data,
      n_pops > 0 & n_pops <= 5
    ),
    aes(
      x=SELECTION_METHOD,
      y=max_tasks_covered,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Maximum task coverage"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette,
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    palette=cb_palette,
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  facet_wrap(
    ~n_pops,
    nrow=1,
    labeller=label_both
  ) +
  theme(
    legend.position="bottom",
    axis.text.x = element_blank()
  )

ggsave(
  paste0(plot_directory, "2021-11-11-task-cov-per-n-pops-alt.pdf"),
  width=10,
  height=6
)
```

## Average number of different populations selected per generation

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_unique_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Avg. number selected"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_brewer(
    name="Selection Method",
    palette=cb_palette,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "2021-11-11-num-selected.pdf")
)
```

```{r}
mean(filter(exp_summary_data, SELECTION_METHOD=="elite")$avg_unique_selected)
mean(filter(exp_summary_data, SELECTION_METHOD=="elite-10")$avg_unique_selected)
mean(filter(exp_summary_data, SELECTION_METHOD=="tourn")$avg_unique_selected)
mean(filter(exp_summary_data, SELECTION_METHOD=="nde")$avg_unique_selected)
mean(filter(exp_summary_data, SELECTION_METHOD=="lex")$avg_unique_selected)
mean(filter(exp_summary_data, SELECTION_METHOD=="random")$avg_unique_selected)
```

### Entropy of selected population IDs

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_entropy_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="none"
  )
```

## Average number of organisms in populations at end of maturation period

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_num_orgs,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Average number of organisms",
    limits=c(950, 1000)
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "2021-11-11-num-orgs.pdf")
)
```

## Average generations per maturation period

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_gens,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Average generations per maturation period"
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_brewer(
    palette=cb_palette
  ) +
  scale_color_brewer(
    palette=cb_palette
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "2021-11-11-avg-gens.pdf")
)

median(exp_summary_data$total_gens_approx) # Used for determining how many generations to run EC for

```

## Manuscript figures

```{r}
grid <- plot_grid(
  max_trait_cov_fig +
    theme(
      axis.title.x=element_blank(),
      axis.text.x = element_text(size = 9)
    ) +
    ggtitle("Best population task coverage"),
  total_trait_cov_fig +
    theme(
      axis.title.x=element_blank(),
      axis.text.x = element_text(size = 9)
    ) +
    ggtitle("Metapopulation task coverage"),
  num_pop_task_profiles_fig +
    theme(
      axis.text.x = element_text(size = 9)
    ) +
    ggtitle("Diversity of task profiles"),
  task_profile_spread_fig +
    theme(
      axis.text.x = element_text(size = 9)
    ) +
    ggtitle("Spread of task profiles"),
  nrow=2,
  ncol=2,
  labels="auto"
)
grid

save_plot(
  filename=paste0(plot_directory, "2021-11-11-selection-figure.pdf"),
  plot=grid,
  base_height=6
)
```