# Population propagule sample size

In this preliminary experiment, we looked at the effect of varying the size of propagules used when creating "offspring" populations from "parent" populations.


We conducted these exploratory experiments well before the final set of experiments presented in our manuscript, so their setups are not the same:

- We only compared elite, lexicase, non-dominated elite, and a no selection control.
- The environment is simpler with 8 population-level functions instead of 18.
- The maximum population size is 900 instead of 1,000.
- The maturation period is longer (300 updates versus 200)
- We ran the experiment for fewer cycles (500 instead of 2,000).

Overall, we found that the effect of propagule size varied by selection scheme.
For elite selection and the no-selection control, sample size had little effect.
For lexicase and non-dominated elite selection, the smallest propagule size (1% of the maximum population size)
resulted in significantly better outcomes than using larger propagule sizes (e.g., 100% of the maximum population size).

Because these data were collected during early experiments, we tracked fewer population/metapopulation statistics.
Future work should further investigate the effects of propagule size, especially in the context of more complex environments that support more complex organism-organism interaction.

## Overview

```{r}
experiment_slug <- "2021-09-30-sample-size"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(scales)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

# Mark factors
exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tournament",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

exp_summary_data$NUM_POPS <- factor(
  exp_summary_data$NUM_POPS,
  levels=c(
    "24",
    "48",
    "96"
  )
)

exp_summary_data$UPDATES_PER_EPOCH <- as.factor(
  exp_summary_data$UPDATES_PER_EPOCH
)

exp_summary_data$POPULATION_SAMPLING_SIZE <- as.factor(
  exp_summary_data$POPULATION_SAMPLING_SIZE
)

exp_summary_data$SAMPLE_SIZE <- exp_summary_data$POPULATION_SAMPLING_SIZE

exp_summary_data <- filter(exp_summary_data, UPDATES_PER_EPOCH=="300")
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Accent"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)

sel.labs <- c(
  "elite",
  "tournament",
  "lex",
  "nde",
  "ndt",
  "random",
  "none"
)
names(sel.labs) <- c(
  "elite",
  "tournament",
  "lex",
  "nde",
  "ndt",
  "random",
  "none"
)
upe.labs <- c(
  "updates per cycle=100",
  "updates per cycle=300"
)
names(upe.labs) <- c(
  "100",
  "300"
)
```

## Average number of organisms

Average number of organisms per world at the end of a run.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=POPULATION_SAMPLING_SIZE,
      y=avg_num_orgs,
      fill=POPULATION_SAMPLING_SIZE
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=POPULATION_SAMPLING_SIZE),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    UPDATES_PER_EPOCH~SELECTION_METHOD,
    labeller = labeller(UPDATES_PER_EPOCH=upe.labs, SELECTION_METHOD=sel.labs)
  ) +
  theme(
    legend.position="none",
    axis.text.x=element_text(angle=45,hjust=1)
  )

ggsave(
  paste0(plot_directory, "avg_num_orgs.pdf")
)
```

In general, the smaller propagule sizes are less likely to reach 900 organisms during the maturation period.
However, all final population sizes are within 25 organisms of each, so no substantial differences here.

## Average generations per maturation period

```{r}
ggplot(
  exp_summary_data,
    aes(
      x=POPULATION_SAMPLING_SIZE,
      y=avg_gens,
      fill=POPULATION_SAMPLING_SIZE
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=POPULATION_SAMPLING_SIZE),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_grid(
    UPDATES_PER_EPOCH~SELECTION_METHOD,
    labeller = labeller(UPDATES_PER_EPOCH=upe.labs, SELECTION_METHOD=sel.labs)
  ) +
  theme(
    legend.position="none",
    axis.text.x=element_text(angle=45,hjust=1)
  )

ggsave(
  paste0(plot_directory, "avg_gens.pdf"),
  width=10,
  height=15
)
```

## Performance

### Best population task coverage

```{r}
ggplot(
  exp_summary_data,
    aes(
      x=POPULATION_SAMPLING_SIZE,
      y=max_trait_coverage,
      fill=POPULATION_SAMPLING_SIZE
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=POPULATION_SAMPLING_SIZE),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage"
  ) +
  facet_grid(
    UPDATES_PER_EPOCH~SELECTION_METHOD,
    labeller = labeller(UPDATES_PER_EPOCH=upe.labs, SELECTION_METHOD=sel.labs)
  ) +
  ggtitle("Best population task coverage") +
  theme(
    legend.position="none",
    axis.text.x=element_text(angle=45,hjust=1),
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_trait_coverage.png"),
  width=10,
  height=6
)
```

```{r}
comp_data <- filter(
  exp_summary_data,
  SELECTION_METHOD=="lex"
)
kruskal.test(
  formula=max_trait_coverage~POPULATION_SAMPLING_SIZE,
  data=comp_data
)

pairwise.wilcox.test(
  x=comp_data$max_trait_coverage,
  g=comp_data$POPULATION_SAMPLING_SIZE,
  p.adjust.method="bonferroni",
  exact=FALSE
)
```

### Metapopulation task coverage

```{r}
ggplot(
  exp_summary_data,
    aes(
      x=POPULATION_SAMPLING_SIZE,
      y=total_trait_coverage,
      fill=POPULATION_SAMPLING_SIZE
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=POPULATION_SAMPLING_SIZE),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage"
  ) +
  facet_grid(
    UPDATES_PER_EPOCH~SELECTION_METHOD,
    labeller = labeller(UPDATES_PER_EPOCH=upe.labs, SELECTION_METHOD=sel.labs)
  ) +
  ggtitle("Metapopulation task coverage") +
  theme(
    legend.position="none",
    axis.text.x=element_text(angle=45,hjust=1),
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "total_trait_coverage.png"),
  width=10,
  height=6
)
```

```{r}
comp_data <- filter(
  exp_summary_data,
  SELECTION_METHOD=="lex"
)
kruskal.test(
  formula=total_trait_coverage~POPULATION_SAMPLING_SIZE,
  data=comp_data
)

pairwise.wilcox.test(
  x=comp_data$total_trait_coverage,
  g=comp_data$POPULATION_SAMPLING_SIZE,
  p.adjust.method="bonferroni",
  exact=FALSE
)
```







