# Varying population maturation period

## Overview

How robust are our digital directed evolution results at different maturation periods?
Adjusting the length of population maturation periods shifts the balance of individual-level and population-level selection pressure.
Shorter maturation periods have shorter stints where organisms compete within a population between population-level selection events;
whereas, longer maturation periods have longer stints where organisms are competing for space within a population before a population-level selection event occurs.

Naively, we might maximize selection pressure on population-level functions by minimizing maturation periods; however, maturation periods of organism-level evolution are important for accumulating genetic variation in a population.
We could maximize genetic divergence among populations in a metapopulation by using long maturation periods; however, if population-level selection events are too infrequent, there may not be sufficient selection pressure to maintain population-level functions if they are not also sufficiently beneficial at the individual-level.
We expect there to be an ideal maturation period length that will vary depending on the particular organisms and functions being evolved.

We ran a supplemental experiment to evaluate the robustness of our results at a range of maturation period lengths.
For each of elite selection, lexicase selection, non-dominated elite selection, and no selection (control), we ran 50 replicates of digital directed evolution with the following maturation period lengths (given in updates):
20,
50,
100,
200  (the default used in our main experiments),
500,
and 1,000.
For each condition, we held the total number of updates constant (400,000 total updates);
that is, we ran fewer overall cycles of directed evolution at larger maturation periods (e.g., 400 cycles for 1,000 update maturation periods), and a greater number of cycles at smaller maturation periods (e.g., 20,000 cycles for 20 update maturation periods).
We used a more limited set of selection schemes for this supplemental experiment due to constraints on computing resources.

```{r}
experiment_slug <- "2021-11-12-time"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(scales)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

exp_summary_data$UPDATES_PER_EPOCH <- as.factor(
  exp_summary_data$UPDATES_PER_EPOCH
)

exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE <- as.factor(
  exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE
)

exp_summary_data$U_PER_E <- exp_summary_data$UPDATES_PER_EPOCH
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Set2"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)

selection_method_breaks <- c("elite", "tourn", "lex", "nde", "random", "none")
selection_method_labels <- c("ELITE", "TOURN", "LEX", "NDE", "RAND", "NONE")

exp_summary_data$SELECTION_METHOD_LABEL <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  ),
  labels=c(
    "Elite Selection",
    "Tournament Selection",
    "Lexicase Selection",
    "Non-dominated Elite Selection",
    "Non-dominated Tournament Selection",
    "Random Selection",
    "No Selection"
  )
)
# 1, 4, 5, 7
selection_methods_smaller_set_colors <- c("#66C2A5", "#E78AC3", "#A6D854", "#E5C494")
# c("#66C2A5", "#8DA0CB", "#E78AC3", "#FFD92F")
```

## Average number of organisms

Average number of organisms per world at the end of a run.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_num_orgs,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="free_y"
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )

ggsave(
  paste0(plot_directory, "avg_num_orgs.pdf")
)
```

```{r}
for (timing in levels(exp_summary_data$U_PER_E)) {
  for (sel in c("elite", "lex", "nde", "none")) {
    print(
      paste(
        timing,
        sel,
        median(filter(exp_summary_data, U_PER_E==timing & SELECTION_METHOD==sel)$avg_num_orgs)
      )
    )
  }
  print(
    paste(
      timing, "overall", median(filter(exp_summary_data, U_PER_E==timing)$avg_num_orgs)
    )
  )
}
```

## Average generations elapsed during the maturation period

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_gens,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  xlab("Updates per maturation period") +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )

ggsave(
  paste0(plot_directory, "avg_gens_facet_sel.pdf"),
  width=20,
  height=10
)
```

## Total generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=total_gens_approx,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )

ggsave(
  paste0(plot_directory, "total_gens_approx.pdf"),
  width=20,
  height=10
)

median(exp_summary_data$total_gens_approx) # Used for determining how many generations to run EC for
```

```{r}
for (timing in levels(exp_summary_data$U_PER_E)) {
  for (sel in c("elite", "lex", "nde", "none")) {
    print(
      paste(
        timing,
        sel,
        median(filter(exp_summary_data, U_PER_E==timing & SELECTION_METHOD==sel)$total_gens_approx)
      )
    )
  }
  print(
    paste(
      timing, "overall", median(filter(exp_summary_data, U_PER_E==timing)$total_gens_approx)
    )
  )
}
```

## Performance

### CPU cycles per replication

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_cpu_cycles_per_replication,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )

ggsave(
  paste0(plot_directory, "avg_cpu_cycles_per_replication.pdf"),
  width=20,
  height=10
)
```

### Best-population task coverage

```{r}

max_task_coverage_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=max_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5, 18.5),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
max_task_coverage_fig
ggsave(
  plot=max_task_coverage_fig,
  filename=paste0(plot_directory, "max_trait_coverage.pdf"),
  width=20,
  height=10
)
```
```{r}
for (sel in c("elite", "lex", "nde", "none")) {
  print(sel)
  sel_summary_data <- filter(exp_summary_data, SELECTION_METHOD==sel)
  kt <- kruskal.test(
    formula=max_trait_coverage~U_PER_E,
    data=sel_summary_data
  )
  print(kt)
  if (kt$p.value < 0.05) {
    pwt <- pairwise.wilcox.test(
      x=sel_summary_data$max_trait_coverage,
      g=sel_summary_data$U_PER_E,
      p.adjust.method="bonferroni",
    )
    print(pwt)
  } else {
    print("Not significant")
  }
}
```

### Metapopulation task coverage

```{r}
total_task_coverage_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=total_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5, 18.5),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
total_task_coverage_fig
ggsave(
  plot=total_task_coverage_fig,
  filename=paste0(plot_directory, "total_trait_coverage.pdf"),
  width=20,
  height=10
)
```

## Population-level Task Profile Diversity

### Task profile richness

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=num_pop_trait_profiles,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  xlab("Maturation period (updates)") +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
ggsave(
  paste0(plot_directory, "num_pop_trait_profiles_facet_sel.pdf"),
  width=20,
  height=10
)
```

### Task profile entropy

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=pop_trait_profile_entropy,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  xlab("Maturation period (updates)") +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
ggsave(
  paste0(plot_directory, "pop_trait_profile_entropy.pdf"),
  width=20,
  height=10
)
```

### Spread (average cosine distance)

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_cosine_dist_from_centroid,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  xlab("Maturation period (updates)") +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
ggsave(
  paste0(plot_directory, "avg_cosine_dist_from_centroid_facet_sel.pdf"),
  width=20,
  height=10
)
```

## Selection

### Average number of unique populations selected

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_unique_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )
ggsave(
  paste0(plot_directory, "avg_unique_selected_facet_sel.pdf"),
  width=20,
  height=10
)
```

### Average entropy of selection ids

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_entropy_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    labeller=label_both
  ) +
  theme(
    legend.position="none",
    axis.text.x = element_text(size = 9, angle=-30, hjust=0)
  )

ggsave(
  paste0(plot_directory, "avg_entropy_selected.pdf")
)
```

## Manuscript Figures

```{r}
grid <- plot_grid(
  max_task_coverage_fig +
    theme(
      axis.text.x = element_text(size = 9, angle=-30, hjust=0),
      strip.text.x = element_text(size = 10)

    ) +
    ggtitle("Best population task coverage"),
  total_task_coverage_fig +
    theme(
      axis.text.x = element_text(size = 9, angle=-30, hjust=0),
      strip.text.x = element_text(size = 10)
    ) +
    ggtitle("Metapopulation task coverage"),
  nrow=2,
  ncol=1,
  labels="auto"
)
grid

save_plot(
  plot=grid,
  filename=paste0(plot_directory, "2021-11-12-timing-figure.pdf"),
  base_height=6
)
```

## Discussion

Lengthening maturation period improves task coverage in the no selection control: there is more genetic variation that can build up between selection bottlenecks, which allows more population-level tasks to evolve by chance.
Performance was surprisingly stable in elite and non-dominated elite selection.
However, lengthening maturation period harms lexicase selection's performance on metapopulation coverage.
Overall, this experiment indicates that different selection schemes may respond differently to adjustments to the balance between individual-level and population-level selection, warranting further exploration in future work.

