# 2021-11-12 - Selection

## Overview

```{r}
experiment_slug <- "2021-11-12-time"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(scales)
library(viridis)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

exp_summary_data$UPDATES_PER_EPOCH <- as.factor(
  exp_summary_data$UPDATES_PER_EPOCH
)

exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE <- as.factor(
  exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE
)

exp_summary_data$U_PER_E <- exp_summary_data$UPDATES_PER_EPOCH
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Set2"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)

selection_method_breaks <- c("elite", "tourn", "lex", "nde", "random", "none")
selection_method_labels <- c("ELITE", "TOURN", "LEX", "NDE", "RAND", "NONE")

exp_summary_data$SELECTION_METHOD_LABEL <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  ),
  labels=c(
    "Elite Selection",
    "Tournament Selection",
    "Lexicase Selection",
    "Non-dominated Elite Selection",
    "Non-dominated Tournament Selection",
    "Random Selection",
    "No Selection"
  )
)
# 1, 4, 5, 7
selection_methods_smaller_set_colors <- c("#66C2A5", "#E78AC3", "#A6D854", "#E5C494")
# c("#66C2A5", "#8DA0CB", "#E78AC3", "#FFD92F")
```

## Average number of organisms

Average number of organisms per world at the end of a run.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=avg_num_orgs,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~U_PER_E,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "avg_num_orgs.pdf")
)
```

```{r}
for (timing in levels(exp_summary_data$U_PER_E)) {
  for (sel in c("elite", "lex", "nde", "none")) {
    print(
      paste(
        timing,
        sel,
        median(filter(exp_summary_data, U_PER_E==timing & SELECTION_METHOD==sel)$avg_num_orgs)
      )
    )
  }
  print(
    paste(
      timing, "overall", median(filter(exp_summary_data, U_PER_E==timing)$avg_num_orgs)
    )
  )
}
```

## Average generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_gens,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  xlab("Updates per epoch") +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "avg_gens_facet_sel.pdf"),
  width=20,
  height=10
)
```

## Total generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=total_gens_approx,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "total_gens_approx.pdf"),
  width=20,
  height=10
)

median(exp_summary_data$total_gens_approx) # Used for determining how many generations to run EC for
```

```{r}
for (timing in levels(exp_summary_data$U_PER_E)) {
  for (sel in c("elite", "lex", "nde", "none")) {
    print(
      paste(
        timing,
        sel,
        median(filter(exp_summary_data, U_PER_E==timing & SELECTION_METHOD==sel)$total_gens_approx)
      )
    )
  }
  print(
    paste(
      timing, "overall", median(filter(exp_summary_data, U_PER_E==timing)$total_gens_approx)
    )
  )
}
```

## Performance

### CPU cycles per replication

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_cpu_cycles_per_replication,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "avg_cpu_cycles_per_replication.pdf"),
  width=20,
  height=10
)
```

### Max trait coverage

```{r}

max_task_coverage_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=max_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5, 18.5),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  plot=max_task_coverage_fig,
  filename=paste0(plot_directory, "max_trait_coverage.pdf"),
  width=20,
  height=10
)
```
```{r}
for (sel in c("elite", "lex", "nde", "none")) {
  print(sel)
  sel_summary_data <- filter(exp_summary_data, SELECTION_METHOD==sel)
  kt <- kruskal.test(
    formula=max_trait_coverage~U_PER_E,
    data=sel_summary_data
  )
  print(kt)
  if (kt$p.value < 0.05) {
    pwt <- pairwise.wilcox.test(
      x=sel_summary_data$max_trait_coverage,
      g=sel_summary_data$U_PER_E,
      p.adjust.method="bonferroni",
    )
    print(pwt)
  } else {
    print("Not significant")
  }
}
```
### Total trait coverage

```{r}
total_task_coverage_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=total_trait_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5, 18.5),
    breaks=seq(0,18,2)
  ) +
  scale_x_discrete(
    name="Maturation period (updates)"
  ) +
  scale_fill_manual(
    values=selection_methods_smaller_set_colors
  ) +
  scale_color_manual(
    values=selection_methods_smaller_set_colors
  ) +
  facet_wrap(
    ~SELECTION_METHOD_LABEL,
    scales="fixed",
    nrow=1
  ) +
  theme(
    legend.position="none"
  )
total_task_coverage_fig
ggsave(
  plot=total_task_coverage_fig,
  filename=paste0(plot_directory, "total_trait_coverage.pdf"),
  width=20,
  height=10
)
```

## Population-level Task Profile Diversity

### Task profile richness

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=num_pop_trait_profiles,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0.1, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  xlab("Updates per epoch") +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "num_pop_trait_profiles_facet_sel.pdf"),
  width=20,
  height=10
)
```

### Task profile entropy

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=pop_trait_profile_entropy,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~U_PER_E,
    scales="free_y",
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "pop_trait_profile_entropy.pdf")
)
```

### Spread (average cosine distance)

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_cosine_dist_from_centroid,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  xlab("Maturation period (updates)") +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "avg_cosine_dist_from_centroid_facet_sel.pdf"),
  width=20,
  height=10
)
```

## Selection

### Average number of unique populations selected

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_unique_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    labeller=label_both,
    nrow=1
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "avg_unique_selected_facet_sel.pdf"),
  width=20,
  height=10
)
```

### Average entropy of selection ids

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=U_PER_E,
      y=avg_entropy_selected,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    scales="fixed",
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "avg_entropy_selected.pdf")
)
```

## Manuscript Figures

```{r}
grid <- plot_grid(
  max_task_coverage_fig +
    theme(
      axis.text.x = element_text(size = 9, angle=-30, hjust=0),
      strip.text.x = element_text(size = 10)

    ) +
    ggtitle("Best population task coverage"),
  total_task_coverage_fig +
    theme(
      axis.text.x = element_text(size = 9, angle=-30, hjust=0),
      strip.text.x = element_text(size = 10)
    ) +
    ggtitle("Metapopulation task coverage"),
  nrow=2,
  ncol=1,
  labels="auto"
)
grid

save_plot(
  plot=grid,
  filename=paste0(plot_directory, "2021-11-12-timing-figure.pdf"),
  base_height=6
)
```

