# Conventional genetic programming experiment

Data analyses for our conventional evolutionary computing experiment.

## Overview

```{r}
experiment_slug <- "2021-11-15-ec"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(khroma)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Load experiment summary data.

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "elite-10",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "elite-10",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)
```

Load time series data.

```{r}
time_series_data_loc <- paste0(working_directory,"data/pop_snapshot_time_series_corrected.csv")
time_series_data <- read.csv(time_series_data_loc, na.strings="NONE")

get_sel <- function(seed) {
  return(filter(exp_summary_data, SEED==seed)$SELECTION_METHOD)
}

solution_evolved_fun <- function(seed, update) {
  d <- filter(exp_summary_data, SEED==seed)
  return(update==d$update && d$max_fit_is_solution=="1");
}

time_series_data$SELECTION_METHOD <- mapply(
  get_sel,
  time_series_data$SEED
)

time_series_data$solution_evolved <- mapply(
  solution_evolved_fun,
  time_series_data$SEED,
  time_series_data$update
)

time_series_data$SELECTION_METHOD <- as.factor(
  time_series_data$SELECTION_METHOD
)

exp_data_gen_2000 <- filter(time_series_data, update==2000)
```

Miscellaneous setup.

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
scale_fill_fun <- scale_fill_bright
scale_color_fun <- scale_color_bright
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)
# Order selection schemes.
selection_method_breaks <- c("elite", "elite-10", "tourn", "lex", "nde", "random", "none")
selection_method_labels <- c("ELITE", "TOP-10", "TOURN", "LEX", "NDE", "RAND", "NONE")
```

## Number of successful replicates

We considered a run to be successful if it produced a program capable of performing all 22 tasks during evaluation.

```{r}
ggplot(
    filter(exp_summary_data, max_fit_is_solution=="1"),
    aes(x=SELECTION_METHOD, fill=SELECTION_METHOD)
  ) +
  geom_bar() +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9),
    vjust=0
  ) +
  scale_y_continuous(
    limits=c(0, 50),
    breaks=seq(0,50,10)
  ) +
  scale_x_discrete(
    name="Selection Method",
    limits=selection_method_breaks,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_fun(
    limits=selection_method_breaks,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_fun(
    limits=selection_method_breaks,
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  ylab("Successful replicates") +
  theme(legend.position = "none")

ggsave(
  paste0(plot_directory, "2021-11-15-num-solutions.pdf")
)
```

## Final task coverage

Task coverage after 55,000 generations of evolution.

```{r}
max_task_cov_fig <-
  ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=max_fit_aggregate_score,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5,22.5),
    breaks=seq(0,22,2)
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_fun(
  ) +
  scale_color_fun(
  ) +
  theme(
    legend.position="none"
    # axis.text = element_text(size = 8),
    # axis.title = element_text(size=10)
  )
max_task_cov_fig

ggsave(
  plot=max_task_cov_fig,
  filename=paste0(plot_directory, "2021-11-15-ec-performance.pdf"),
  height=3,
  width=4
)
```

Statistical results:

```{r}
kruskal.test(
  formula=max_fit_aggregate_score~SELECTION_METHOD,
  data=exp_summary_data
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_summary_data$max_fit_aggregate_score,
  g=exp_summary_data$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

## Generation 2,000 task coverage

Task coverage after 2,000 generations (i.e., the number of cycles runin the directed evolution experiments)

```{r}
ggplot(
    exp_data_gen_2000,
    aes(
      x=SELECTION_METHOD,
      y=max_org_task_coverage,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8,
    adjust=1.5
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5,22.5),
    breaks=seq(0,22,2)
  ) +
  scale_x_discrete(
    name="Selection Method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_fill_fun() +
  scale_color_fun() +
  theme(
    legend.position="none",
    axis.text = element_text(size = 8),
    axis.title = element_text(size=10)
  )

ggsave(
  filename=paste0(plot_directory, "max_aggregate_score_gen_2000.pdf"),
  height=3,
  width=4
)
```

Statistical results:

```{r}
kruskal.test(
  formula=max_org_task_coverage~SELECTION_METHOD,
  data=exp_data_gen_2000
)

# Kruskal-wallis is significant, so we do a post-hoc wilcoxon rank-sum.
pairwise.wilcox.test(
  x=exp_data_gen_2000$max_org_task_coverage,
  g=exp_data_gen_2000$SELECTION_METHOD,
  p.adjust.method="bonferroni",
)
```

## Population-level task coverage

Task coverage across entire population after 55,000 generations of evolution.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=population_num_tasks_covered,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_fun() +
  scale_color_fun() +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "population_num_tasks_covered.pdf")
)
```

## Generations elapsed before a solution evolves

Runs where no solution evolved are in gray and plotted as "unsolved".

```{r}
unfinished_data <- filter(exp_summary_data, max_fit_is_solution=="0")
unfinished_data$graph_update <- 60000

ggplot(
    filter(exp_summary_data, max_fit_is_solution=="1"),
    aes(
      x=SELECTION_METHOD,
      y=update
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_point(
    data = unfinished_data,
    mapping=aes(
      x=SELECTION_METHOD,
      y=graph_update
    ),
    color="gray",
    position = position_jitter(width = .15, height=1000),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    name="Generation first solution evolved",
    limits=c(0, 65000),
    breaks=c(0, 10000, 20000, 30000, 40000, 50000, 60000),
    labels=c("0", "10000", "20000", "30000", "40000", "50000", "Unsolved")
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "updates_until_solution.pdf")
)
```

## Best task coverage over time

```{r}
max_task_cov_ts <-
  ggplot(
    time_series_data,
    aes(
      x=update,
      y=max_org_task_coverage,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    name="Task Coverage",
    limits=c(-0.5,22.5),
    breaks=seq(0,22,2)
  ) +
  scale_x_continuous(
    name="Generation",
    limits=c(0, 55000),
    breaks=c(0, 10000, 20000, 30000, 40000, 50000)
  ) +
  scale_fill_fun(
    name="Selection method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  scale_color_fun(
    name="Selection method",
    breaks=selection_method_breaks,
    labels=selection_method_labels
  ) +
  theme(
    legend.position="none"
  )
max_task_cov_ts

ggsave(
  paste0(plot_directory, "2021-11-15-ec-performance-ts.pdf")
)
```

## Manuscript Figure

```{r}
legend <- cowplot::get_legend(
    max_task_cov_ts +
      guides(
        color=guide_legend(nrow=1),
        fill=guide_legend(nrow=1)
      ) +
      theme(
        legend.position = "bottom",
        legend.box="horizontal",
        legend.justification="center"
      )
  )

grid <- plot_grid(
  max_task_cov_ts +
    ggtitle("Task coverage over time") +
    labs(subtitle="") +
    theme(legend.position="none"),
  max_task_cov_fig +
    ggtitle("Final task coverage") +
    theme(),
  nrow=1,
  ncol=2,
  align="h",
  # rel_widths=c(3,2),
  labels="auto"
)
# grid
grid <- plot_grid(
  grid,
  legend,
  nrow=2,
  ncol=1,
  rel_heights=c(1, 0.1)
)
grid

save_plot(
  paste(
    plot_directory,
    "2021-11-15-performance-fig.pdf",
    sep=""
  ),
  grid,
  base_width=12,
  base_height=6
)
```