# 2021-11-15 - Evolutionary computing setup

## Overview

```{r}
experiment_slug <- "2021-11-15-ec"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(scales)
library(viridis)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)
```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Set2"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)
```

## Performance

### Max aggregate performance

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=max_fit_aggregate_score,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "max_aggregate_score.pdf")
)
```

### Population-level task coverage

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=SELECTION_METHOD,
      y=population_num_tasks_covered,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="none"
  )

ggsave(
  paste0(plot_directory, "population_num_tasks_covered.pdf")
)
```

### Number of solutions

```{r}
ggplot(
    filter(exp_summary_data, max_fit_is_solution=="1"),
    aes(x=SELECTION_METHOD, fill=SELECTION_METHOD)
  ) +
  geom_bar() +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9),
    vjust=0
  ) +
  scale_y_continuous(
    limits=c(0, 50),
    breaks=seq(0,50,10)
  ) +
  scale_x_discrete(
    limits=c(
      "elite",
      "tourn",
      "lex",
      "nde",
      "random",
      "none"
    ),
    breaks=c(
      "elite",
      "tourn",
      "lex",
      "nde",
      "random",
      "none"
    )
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  ylab("Successful replicates") +
  theme(legend.position = "none")

ggsave(
  paste0(plot_directory, "num_solutions.pdf")
)
```

### How many generations elapse before solutions evolve?

```{r}
unfinished_data <- filter(exp_summary_data, max_fit_is_solution=="0")
unfinished_data$graph_update <- 60000

ggplot(
    filter(exp_summary_data, max_fit_is_solution=="1"),
    aes(
      x=SELECTION_METHOD,
      y=update,
      fill=SELECTION_METHOD
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=SELECTION_METHOD),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_point(
    data = unfinished_data,
    mapping=aes(
      x=SELECTION_METHOD,
      y=graph_update
    ),
    color="gray",
    position = position_jitter(width = .15, height=1000),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  scale_y_continuous(
    name="Generation first solution evolved",
    limits=c(0, 65000),
    breaks=c(0, 10000, 20000, 30000, 40000, 50000, 60000),
    labels=c("0", "10000", "20000", "30000", "40000", "50000", "Unsolved")
  ) +
  theme(
    legend.position="none"
  )
ggsave(
  paste0(plot_directory, "updates_until_solution.pdf")
)
```

## Time series data

```{r}
time_series_data_loc <- paste0(working_directory,"data/pop_snapshot_time_series.csv")
time_series_data <- read.csv(time_series_data_loc, na.strings="NONE")

get_sel <- function(seed) {
  return(filter(exp_summary_data, SEED==seed)$SELECTION_METHOD)
}


solution_evolved_fun <- function(seed, update) {
  d <- filter(exp_summary_data, SEED==seed)
  return(update==d$update && d$max_fit_is_solution=="1");
}

time_series_data$SELECTION_METHOD <- mapply(
  get_sel,
  time_series_data$SEED
)

time_series_data$solution_evolved <- mapply(
  solution_evolved_fun,
  time_series_data$SEED,
  time_series_data$update
)

time_series_data$SELECTION_METHOD <- as.factor(
  time_series_data$SELECTION_METHOD
)

# Specify experimental condition for each datum.
# times_series_data$SELECTION_METHOD <- factor(
#   times_series_data$SELECTION_METHOD,
#   levels=c(
#     "elite",
#     "tournament",
#     "lexicase",
#     "non-dominated-elite",
#     "non-dominated-tournament",
#     "random",
#     "none"
#   ),
#   labels=c(
#     "elite",
#     "tourn",
#     "lex",
#     "nde",
#     "ndt",
#     "random",
#     "none"
#   )
# )
```

Note that these time series plots are a little weird in that it only shows data from runs where solutions have not evolved yet.
Once a solution evolves, we stopped the run, so no data exists.

### Population-level task coverage

```{r}
ggplot(
    filter(time_series_data, solution_evolved==FALSE),
    aes(
      x=update,
      y=population_task_coverage,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  paste0(plot_directory, "population_task_coverage_ot.pdf"),
  width=20,
  height=10
)
```

### Max organism task coverage

max_org_task_coverage
```{r}
ggplot(
    filter(time_series_data, solution_evolved==FALSE),
    aes(
      x=update,
      y=max_org_task_coverage,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  paste0(plot_directory, "max_org_task_coverage_ot.pdf"),
  width=20,
  height=10
)
```


### Number of unique task profiles

```{r}
ggplot(
    filter(time_series_data, solution_evolved==FALSE),
    aes(
      x=update,
      y=num_unique_task_profiles,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  paste0(plot_directory, "num_unique_task_profiles_ot.pdf"),
  width=20,
  height=10
)
```

### Task profile spread

(avg_cosine_dist_from_centroid)

```{r}
ggplot(
    filter(time_series_data, solution_evolved==FALSE),
    aes(
      x=update,
      y=avg_cosine_dist_from_centroid,
      fill=SELECTION_METHOD,
      color=SELECTION_METHOD
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_fill_brewer(
    palette="Set2"
  ) +
  scale_color_brewer(
    palette="Set2"
  ) +
  theme(
    legend.position="bottom"
  )

ggsave(
  paste0(plot_directory, "avg_cosine_dist_from_centroid_ot.pdf"),
  width=20,
  height=10
)
```