# 2021-11-08 - Longer genomes

## Overview

```{r}
experiment_slug <- "2021-11-08-longer-genomes"

working_directory <- paste0("experiments/",experiment_slug,"/analysis/")
```

## Analysis dependencies

Load all required R libraries

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggpcp)
library(cowplot)
library(RColorBrewer)
library(scales)
library(viridis)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were knit with the following environment:

```{r}
print(version)
```

## Setup

Experiment summary data

```{r}
exp_summary_data_loc <- paste0(working_directory,"data/experiment_summary.csv")
exp_summary_data <- read.csv(exp_summary_data_loc, na.strings="NONE")

exp_summary_data$SELECTION_METHOD <- factor(
  exp_summary_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

exp_summary_data$AVIDAGP_ENV_FILE <- factor(
  exp_summary_data$AVIDAGP_ENV_FILE,
  levels=c(
    "environment-no-indiv.json",
    "environment-big.json"
  ),
  labels=c(
    "env-bb-0",
    "env-bb-1"
  )
)


exp_summary_data$NUM_POPS <- factor(
  exp_summary_data$NUM_POPS,
  levels=c(
    "24",
    "48",
    "96"
  )
)

exp_summary_data$UPDATES_PER_EPOCH <- as.factor(
  exp_summary_data$UPDATES_PER_EPOCH
)

exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE <- as.factor(
  exp_summary_data$TOURNAMENT_SEL_TOURN_SIZE
)

exp_summary_data$POPULATION_SAMPLING_SIZE <- as.factor(
  exp_summary_data$POPULATION_SAMPLING_SIZE
)

exp_summary_data$genome_length <- factor(
  exp_summary_data$ANCESTOR_FILE,
  levels=c(
    "ancestor-50.gen",
    "ancestor-100.gen",
    "ancestor-150.gen",
    "ancestor-200.gen"
  ),
  labels=c(
    "50",
    "100",
    "150",
    "200"
  )
)

exp_summary_data$SAMPLE_SIZE <- exp_summary_data$POPULATION_SAMPLING_SIZE
exp_summary_data$ENV <- exp_summary_data$AVIDAGP_ENV_FILE
exp_summary_data$U_PER_E <- exp_summary_data$UPDATES_PER_EPOCH


```

Miscellaneous setup

```{r}
# Configure our default graphing theme
theme_set(theme_cowplot())
# Palette
cb_palette <- "Paired"
# Create a directory to store plots
plot_directory <- paste0(working_directory, "plots/")
dir.create(plot_directory, showWarnings=FALSE)
```

## Average number of organisms

Average number of organisms per world at the end of a run.

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_num_orgs,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_num_orgs.pdf")
)
```

## Average generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_gens,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_gens.pdf"),
  width=10,
  height=15
)
```

## Total generations

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=total_gens_approx,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "total_gens_approx.pdf"),
  width=10,
  height=15
)
```

## Performance

### CPU cycles per replication

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_cpu_cycles_per_replication,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_cpu_cycles_per_replication.pdf")
)
```

### Max aggregate performance

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=max_aggregate_score,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  # scale_y_continuous(
  #   name="max_aggregate_score (log scale)",
  #   trans=pseudo_log_trans(sigma = 1, base = 10)
  # ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_aggregate_score.pdf")
)
```

### Max trait coverage

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=max_trait_coverage,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_trait_coverage.pdf")
)
```

### Total trait coverage

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=total_trait_coverage,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "total_trait_coverage.pdf")
)
```

## Population-level Task Profile Diversity

### Task profile richness

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=num_pop_trait_profiles,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "num_pop_trait_profiles.pdf")
)
```

### Task profile entropy

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=pop_trait_profile_entropy,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "pop_trait_profile_entropy.pdf")
)
```

### Spread (avg cosine distance)

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_cosine_dist_from_centroid,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_cosine_dist_from_centroid.pdf")
)
```

## Selection

### Average number of unique populations selected

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_unique_selected,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_unique_selected.pdf")
)
```

### Average entropy of selection ids

```{r}
ggplot(
    exp_summary_data,
    aes(
      x=genome_length,
      y=avg_entropy_selected,
      fill=genome_length
    )
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=genome_length),
    position = position_jitter(height=0, width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  facet_wrap(
    ~SELECTION_METHOD
  ) +
  theme(
    legend.position="none",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "avg_entropy_selected.pdf")
)
```

## Time series

Load data

```{r}
times_series_data_loc <- paste0(working_directory,"data/evaluation_time_series.csv")
times_series_data <- read.csv(times_series_data_loc, na.strings="NONE")

get_genome_length <- function(seed) {
  return(filter(exp_summary_data, SEED==seed)$genome_length)
}

# Specify experimental condition for each datum.
times_series_data$SELECTION_METHOD <- factor(
  times_series_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

times_series_data$AVIDAGP_ENV_FILE <- factor(
  times_series_data$AVIDAGP_ENV_FILE,
  levels=c(
    "environment-no-indiv.json",
    "environment-big.json"
  ),
  labels=c(
    "env-bb-0",
    "env-bb-1"
  )
)

times_series_data$NUM_POPS <- factor(
  times_series_data$NUM_POPS,
  levels=c(
    "24",
    "48",
    "96"
  )
)

times_series_data$UPDATES_PER_EPOCH <- as.factor(
  times_series_data$UPDATES_PER_EPOCH
)

times_series_data$genome_length <- mapply(
  get_genome_length,
  times_series_data$SEED
)

times_series_data$U_PER_E <- times_series_data$UPDATES_PER_EPOCH
```

### Max trait coverage

```{r}
ggplot(
    times_series_data,
    aes(
      x=updates_elapsed,
      y=max_trait_coverage,
      fill=genome_length,
      color=genome_length
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    nrow=2
  ) +
  theme(
    legend.position="bottom",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "max_trait_coverage_ot.png"),
  width=20,
  height=10
)
```

```{r}
ggplot(
    times_series_data,
    aes(
      x=updates_elapsed,
      y=total_trait_coverage,
      fill=genome_length,
      color=genome_length
    )
  ) +
  stat_summary(geom="line", fun=mean) +
  stat_summary(
    geom="ribbon",
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    alpha=0.2,
    linetype=0
  ) +
  scale_y_continuous(
    breaks=seq(0,18,2)
  ) +
  facet_wrap(
    ~SELECTION_METHOD,
    nrow=2
  ) +
  theme(
    legend.position="bottom",
    panel.border=element_rect(colour="grey",size=1)
  )

ggsave(
  paste0(plot_directory, "total_trait_coverage_ot.png"),
  width=20,
  height=10
)
```

## Pairwise Population comparisons

```{r}
pairwise_pop_data_loc <- paste0(working_directory,"data/pairwise_population_comps.csv")
pairwise_pop_data <- read.csv(pairwise_pop_data_loc, na.strings="NONE")

# Specify experimental condition for each datum.
pairwise_pop_data$SELECTION_METHOD <- factor(
  pairwise_pop_data$SELECTION_METHOD,
  levels=c(
    "elite",
    "tournament",
    "lexicase",
    "non-dominated-elite",
    "non-dominated-tournament",
    "random",
    "none"
  ),
  labels=c(
    "elite",
    "tourn",
    "lex",
    "nde",
    "ndt",
    "random",
    "none"
  )
)

pairwise_pop_data$genome_length <- factor(
  pairwise_pop_data$ANCESTOR_FILE,
  levels=c(
    "ancestor-50.gen",
    "ancestor-100.gen",
    "ancestor-150.gen",
    "ancestor-200.gen"
  ),
  labels=c(
    "50",
    "100",
    "150",
    "200"
  )
)

```

```{r}
# 60224
ggplot(
    filter(pairwise_pop_data, SEED=="60224"),
    aes(x=pop_a, y=pop_b, fill=cosine_distance)
  ) +
  geom_tile(
    color="white",
    size=0.1,
    width=1,
    height=1
  ) +
  # geom_tile(
  #   data=filter(profile_data, pop_level=="1" & task_coverage=="0"),
  #   fill="#666666",
  #   color="white",
  #   size=0.1,
  #   width=1,
  #   height=1
  # ) +
  scale_fill_viridis(discrete=FALSE) +
  theme(
    axis.text.x=element_text(angle=45, hjust=1)
  )

```

